# IE - Master in Business Analytics and Big Data
Recommendation Systems

Building a Collaborative Filtering Engine

Author: Kamal Nandan

========================================================

# Initialization
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import libraries
```{r}
library(data.table)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(grid)
```

Load the critics dataset.
```{r}
rdf <- data.frame(read.csv("critics.csv", header = TRUE, stringsAsFactors = FALSE))
#rdf1 = read.csv("critics.csv")
```

# 1) Pearson correlation coefficient 
Calculate the Pearson correlation coefficient between Victoria and Nuria

**Expected correlation between Victoria and Nuria**: 0.3651484

```{r fig.width=7, fig.height=6}
ratingV <- as.numeric(as.vector(rdf[rdf$User=="Victoria", 2:21])) # Extract Victoria's ratings
ratingN <- as.numeric(as.vector(rdf[rdf$User=="Nuria", 2:21])) # Extract Nuria's ratings
r1 <- cor(ratingV, ratingN, method = c("pearson"), use = "pairwise.complete.obs") # Correlation coefficient
paste("Correlation between Victoria and Nuria is:", r1)

```

# 2) Compare two movie critics
Using the calculated Pearson correlation in the previous section. Compare and show the movie ratings for Victoria and Nuria on a scatter plot. How is this compared with the same ratings of Maria and Nerea on a scatter plot? What about Chris and Jim? What it means?
```{r}
rating_Victoria <- as.numeric(as.vector(rdf[rdf$User=="Victoria", 2:21]))
rating_Nuria <- as.numeric(as.vector(rdf[rdf$User=="Nuria", 2:21]))
p1 = qplot(rating_Victoria, rating_Nuria, geom = c("point","smooth"))
r1 <- cor(rating_Victoria, rating_Nuria, method = c("pearson"), use = "pairwise.complete.obs")
paste("Correlation between Victoria and Nuria is:", r1)


rating_Maria <- as.numeric(as.vector(rdf[rdf$User=="Maria", 2:21]))
rating_Nerea <- as.numeric(as.vector(rdf[rdf$User=="Nerea", 2:21]))
p2 = qplot(rating_Maria, rating_Nerea, geom= c("point","smooth"))
r1 <- cor(rating_Maria, rating_Nerea, method = c("pearson"), use = "pairwise.complete.obs")
paste("Correlation between Maria and Nerea is:", r1)


rating_Chris <- as.numeric(as.vector(rdf[rdf$User=="Chris", 2:21]))
rating_Jim <- as.numeric(as.vector(rdf[rdf$User=="Jim", 2:21]))
p3 = qplot(rating_Chris, rating_Jim, geom=c("point","smooth"))
r1 <- cor(rating_Chris, rating_Jim, method = c("pearson"), use = "pairwise.complete.obs")
paste("Correlation between Chris and Jim is:", r1)


grid.arrange(p1, p2, p3, nrow = 1)

# Looking at the graph and correlation figures we can say that:
# There is slightly positive correlation(0.36) between Victoria and Nuria
# There is a slightly negative correlation(-0.3) between Maria and Nerea
# There is a high correlation between (0.825) between Chris and Jim

```


# 3) Top Recommendations 
Get the top 5 movies recommended for Victoria. by using a weighted average of every other critic's rankings.

**Expected result**:
3.7917013044215, 'The Matrix'
3.50776533175371, 'Forrest Gump'     
3.33118834864677, 'The Sixth Sense' 
3.11491825315719, 'Shakespeare in Love'
2.9124513228665, 'Blade Runner'

```{r fig.width=7, fig.height=6}
trdf <- t(rdf)            ## Traspose the data frame
colnames(trdf) = trdf[1,]  ## Assign the column names
trdf =trdf[-1,]            ## Eliminate first row

## Convert it to data frame in order to transform values to numeric
d2 <- data.frame(trdf) 
d2[1:20] <- lapply(d2[1:20], function(x) as.numeric(as.character(x)))
d2.1=d2[1:20]

# Keep out movies that Victoria hasn't watched
d2.v=d2.1[complete.cases(d2.1$Victoria != "NA"),,drop=F]

 ## Correlation table
d3=cor(d2.1[,unlist(lapply(d2.1,is.numeric))],use="pairwise.complete.obs")

## Select the movies that Victoria hasn't watch yet
MOVIES=rownames(d2.1[is.na(d2.1$Victoria == "NA"),15,drop=F])

## User ratings for all the movies that Victoria hasn't watch yet
dff=rdf[1:20,c(MOVIES)]

## Calculate the mean for all the users
user_mean = rowMeans(rdf[,2:21],na.rm=T)

## Calculate the difference between the Rating of each movie with its respective User Mean
almostdiff=cbind(dff,user_mean)
diff = (almostdiff-almostdiff$user_mean)[-15,-8]

## Multiply the previous Difference by each user Correlation that Victoria.
almostdiff2=cbind(diff,corr=(d3[,15])[-15])
df3=colSums((almostdiff2*almostdiff2$corr)[-8],na.rm=T)

## Divide it by the Sum of all the Pearson correlations
df3=df3/sum(abs((d3[,15])[-15]),na.rm=T)

## Finally sum all the movie results with Victoria rating's mean
df3=df3+ mean(d2.v$Victoria)
head(sort(df3,decreasing=T),n=5)

```

# 4) Top similar critics
Return the top 5 critics that are most similar to Victoria. 

**Expected result**:
0.9449112, 'Rachel'
0.5976143, 'Ana' 
0.5789794, 'Oriol'
0.4925922, 'Maria'
0.4273247, 'Carles'

```{r fig.width=7, fig.height=6}

ratingVictoria <- as.numeric(as.vector(rdf[rdf$User=="Victoria", 2:21])) # Extract Victoria's ratings
user_names <- rdf[rdf$User!="Victoria", 1] # Get the names of all the usrers other than Victoria
corr_vector = vector(mode="integer", length=20)

# Loop throw the dataframe and find the correlation between Victoria and everyone else
for(i in 1:nrow(rdf)) {
  if(rdf[i,1]=="Victoria") {
    corr_vector[i] = -1
  } else {
    userrating <- as.numeric(as.vector(rdf[i, 2:21]))
    corr_vector[i] = cor(ratingVictoria, userrating, method = c("pearson"), use = "pairwise.complete.obs")
  }
}

# Add a new column that will contain the correlation between Victoria and everyone else
victoriaVsOthers <- rdf %>% mutate(victoriacorr = corr_vector)  

#Arrange the dataframe in descending order based on correlation between Victoria and everyone else
trdf_d_by_pcgt4 <- victoriaVsOthers %>% select(User, victoriacorr) %>% arrange(desc(victoriaVsOthers$victoriacorr))

# print the top 5 correlated users
print(trdf_d_by_pcgt4[1:5,])

```