# IE - Master in Business Analytics and Big Data
Recommendation Systems

Building a Non Personalized Recommendation Engine

Author: FirstName LastName

========================================================

# Initialization
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import libraries
```{r}
library(data.table)
library(dplyr)
library(plyr)
```


```{r}
#function to transpose a dataframe, if required
transpose_data_frame <- function(df) {
  t_df <- transpose(df)
  colnames(t_df) <- rownames(df)
  rownames(t_df) <- colnames(df)
  return (t_df)
}
```

Load the critics dataset. Also do some data cleaning or transformation

```{r}
rdf <- data.frame(read.csv("critics.csv", header = TRUE, stringsAsFactors = FALSE))

mnames <- colnames(rdf) # contain the movie names
mnames <- mnames[-1] #first one is not a movie name
unames <- rdf[,1] # contain the usernames

#remove the first column from the data frame - thats useless
rdf1 = rdf
rdf <- rdf[, -1]

avgrating <- colMeans(rdf, na.rm = TRUE)
# append average rating to the data frame as a row
rdfwithmean <- rbind(rdf, avgrating)

#transpose the data frame now
trdf <- transpose_data_frame(rdfwithmean)

trdf <- trdf %>% mutate(Movie = mnames) # we can also use "trdf <- cbind(trdf, mnames)" or perhaps (trdf$Movie <- mnames)
```

# 1) Top 5 Movies, by rating mean
Calculate mean rating for each movie, ordered with the highest rating listed first, and plot the top 5.

**Expected result**:
3.600000, 'The Shawshank Redemption'
3.266667, 'Star Wars IV - A New Hope'
3.222222, 'Blade Runner'
3.166667, 'Groundhog Day'
3.062500, 'The Silence of the Lambs'

```{r Top 5 Movies, by rating mean}
trdfd <- trdf %>% select(`21`, Movie) %>% arrange(desc(trdf$`21`))
print(trdfd[1:5,])
```

# 2) Top 5 Movies, by rating distribution
Calculate the percentage of ratings for each movie that are 4 stars or higher. Order with the highest percentage first, and plot the top 5 movies with more high ratings distribution.

**Expected result**:
0.7000000, 'The Shawshank Redemption'
0.5333333, 'Star Wars IV - A New Hope'
0.5000000, 'Gladiator'
0.4444444, 'Blade Runner'
0.4375000, 'The Silence of the Lambs'

```{r fig.width=7, fig.height=6}
get_pc_4_or_5 <- function(ratings1) {
  t <- table(ratings1)
  return (sum(t[names(t) >= 4]) / sum(!is.na(ratings1)))
} 

numofmovies = nrow(trdf)
pcmorethan4 = vector(mode = "double", length = numofmovies)
for (row in 1:numofmovies) {
  ratings = trdf[row, 1:20]
  pcmorethan4[row] <- get_pc_4_or_5(unlist(ratings))
}

print(pcmorethan4)
trdf <- trdf %>% mutate("PCgt4" = pcmorethan4)

trdf_d_by_pcgt4 <- trdf %>% select(PCgt4, Movie) %>% arrange(desc(trdf$PCgt4))
print(trdf_d_by_pcgt4[1:5,])

```

# 3) Top 5 Movies, by quantity of ratings
Count the number of ratings for each movie, order the one with most number of ratings first, submit the top 5.

**Expected result**:
17, 'Toy Story'
16, 'The Silence of the Lambs'
15, 'Star Wars IV - A New Hope'
14, 'Star Wars VI - Return of the Jedi'
13, 'Independence Day'

```{r fig.width=7, fig.height=6}
numofmovies = nrow(trdf)
no_of_ratings = vector(mode = "integer", length = numofmovies)
for (row in 1:numofmovies) {
  ratings = trdf[row, 1:20]
  no_of_ratings[row] <- sum(!is.na(ratings))
}

print(no_of_ratings)
trdf <- trdf %>% mutate("NoOfRatings" = no_of_ratings)

trdf_d_by_NoOfRtgs <- trdf %>% select(NoOfRatings, Movie) %>% arrange(desc(trdf$NoOfRatings))
print(trdf_d_by_NoOfRtgs[1:5,])
```

# 4) People who watched Star Wars IV also watched ...
Calculate movies that most often occur with other movie. For each movie, calculate the percentage of the other movie raters who also rated that movie. Order with the highest percentage first, and submit the top 5. Select "Star.Wars.IV...A.New.Hope" movie by defaut. Use the (x+y)/x method.

Top Occurences:
0.9333333, 'Toy Story'
0.8666667, 'Star Wars: Episode VI - Return of the Jedi'
0.8000000, 'The Silence of the Lambs'
0.7333333, 'Independence Day'
0.6666667, 'Total Recall'

```{r}
#rdf1=read.csv("critics.csv")
top5 = rdf1[complete.cases(rdf1$Star.Wars.IV...A.New.Hope), ] #We identify the complete rated critics of the movies
top5 = top5[2:20] != "NA"  #We exclude the NAs from the data, so the movies which are not rated
#View(cm5)
top5.1 = colSums(top5, na.rm = T) # We sum up the rating of each movie
top5.1 = sort(top5.1, decreasing = T) # We sort in descending order
top5.1 = head(top5.1, n = 6) # We finally select the top 6 movies rating, 6 because we need to keep as well "Star Wars IV...a New Hope"
top5.1 = data.frame(common=head(top5.1)) # The Data are transformed into a data frame
#View(cm5.1) # Let's see the rating of the top 6 movie rating
cm6 = top5.1[2:6,,drop=F]/(top5.1[1,]) # Last step is to calculate the percentage of the other movie raters who also rated that movie
head(cm6, n = 5) # We print the Top Occurences
```


# Ignore the follwoing block of code; This function too calculates the ratings, but using a conventional approach;However this one finds the top5 for each and every movie
```{r fig.width=7, fig.height=6}
# This function too calculates the ratings, but using a conventional approach; Please ignore this. However this one 
# finds the top5 for each and every movie
# create a 20*20 matrix that contains the association confidence between each movies
MY_AND <- function(...) Reduce("&", list(...))

no_of_movies = nrow(trdf)
masoc = matrix(data=c(0), nrow=no_of_movies, ncol=no_of_movies)
no_of_critics = 20  #Find a better way of doing this
for(i in 1:no_of_movies) {
  for(j in 1:no_of_movies) {
    if(i==j) {
      masoc[i,j] = -1
    } else {
      x <- !is.na(trdf[i, 1:no_of_critics])
      y <- !is.na(trdf[j, 1:no_of_critics])
      #masoc[i,j]=(sum(MY_AND(x,y))) / (sum(x))
      masoc[j,i]=(sum(MY_AND(x,y))) / (sum(x))
    }
  }
}

movie_names <- rownames(trdf)
col_names <- c("movie names", movie_names)
assoc_matrix <- masoc
assoc_matrix <- data.frame(cbind(movie_names, assoc_matrix))
colnames(assoc_matrix) <- col_names

get_most_recommended<-function(amatrix, movie_name) {
  #movie_names <- amatrix[, 1]
  #for(movie in movie_names) {
    movie_cor = amatrix[, c("movie names", movie_name)]
    print("=================================\n")
    print(movie_name)
    print("=================================\n")
    #movie_cor_d <- movie_cor %>% arrange(desc(movie_name))
    movie_cor_d <- movie_cor[order(movie_cor[,movie_name], decreasing = TRUE),]
    print(movie_cor_d[1:5,])
  #}
}

movie_names <- assoc_matrix[, 1]
#get_most_recommended(assoc_matrix, "Gladiator")

for(movie in movie_names) {
  get_most_recommended(assoc_matrix, movie)
}

```

# 5) People who liked Babe also liked ...
Calculate the movies better rated by people who liked a movie. Select the people who liked the movie "Babe" (4 or 5 stars) and provide the top 5 movies they liked most.

**Expected result**:
5.000, 'Pulp.Fiction'
5.000, 'Groundhog.Day'
4.500, 'The.Shawshank.Redemption'
4.333, 'Toy.Story'
4.000, 'Blade.Runner'

```{r fig.width=7, fig.height=6}
likedBabeDf=rdf1[complete.cases(rdf$Babe), ] # We select the rated critics from Babe movie
likedBabeDf=likedBabeDf[(likedBabeDf$Babe==4|likedBabeDf$Babe==5),] # We select only critics with rating 4 or 5
#View(likedBabeDf) # Let's view the Data
likedBabeDf1=likedBabeDf[-which(colnames(likedBabeDf)=="Babe")] # At that point we don'T need anymore the Babe rating
likedBabeDf1=likedBabeDf1[-which(colnames(likedBabeDf1)=="User")] # We don't need the name of the rater
#View(likedBabeDf1) # Let's view the data
likedBabeDf2 = colMeans(likedBabeDf1, na.rm=T) # Let's calculate the average rating
likedBabeDf2 = sort(likedBabeDf2, decreasing = T) # We sort in descending order
likedBabeDf2 = head(likedBabeDf2, n = 5) # We select the top 5 movies
head(likedBabeDf2) # We print out the movies better rated of people who liked Babe with 4 or 5 stars.
```

# 6) MovieLens
Explore a real data set and provide non-personalized ratings. You can use the MovieLens database. You can find MovieLens' dataset here: http://files.grouplens.org/datasets/movielens/ml-10m-README.html
